Magnetised Polish doughnuts revisited

### 2017-06-16 paper_revised/referee_reply.txt

Dear A&A Editor,

Thank you for sending us the report of the referee, whom we thank for
carefully reading our manuscript and providing very useful feedback. We 
have made several changes addressing the issues raised in the report, 
which are listed below together with our response. The changes made in 
the manuscript are highlighted in boldface, to facilitate identification. 
We hope that with these amendments the paper will be considered 
acceptable for publication on Astronomy and Astrophysics.

       Best wishes,

       The authors

----------------------------------------------------------------------
Report of the Referee -- AA/2017/30935
----------------------------------------------------------------------

> The paper presents a new procedure to construct sequences of equilibrium
> tori around Kerr black holes that are endowed with a toroidal magnetic
> field. These configurations are not expected to be produced under generic
> conditions but the construction of such equilibria is interesting per se
> and has a rather long history. Furthermore, as discussed by the authors,
> the availability of such equilibria can be employed in test simulations
> of general-relativistic codes in ideal MHD.

> The paper is nicely written with a clear discussion of the procedure
> followed and of the properties of the tori that have been found. The
> paper is essentially ready for publication but there are a few aspects
> that could be improved and that I ask the authors to consider.
>

> - Title: I frankly find it unnecessarily colloquial. I do not insist but
> I think a more pragmatic title such as "A new method for the
> construction of magnetized tori around black holes" is scientifically
> more informative.

We agree with the referee about the title being colloquial. Our choice was
motivated by the title of the paper of Qian et al (2009), “The Polish 
doughnuts revisited …”, of which our paper might be considered an extension 
(by incorporating magnetic fields). We would prefer to keep the original 
title, if the referee concurred. If that were not the case, we would not
object to having it changed to his/her suggestion. (“A new method for the
construction of magnetized tori around black holes”.)

> - Sec. 2.1: the authors introduce the concept of cusp but do not really
> define it

> - Sec. 2.1: they introduce the equation of state (13) but do not discuss
> how they treat the specific internal energy that is part of the
> enthalpy used in (13). A comment should be added.
>

> - Sec. 2.1: I am not clear how the magnetic field is actually
> prescribed. The author mention that the only nonzero component of
> the magnetic field in the comoving frame is the azimuthal one but I am
> still unclear how the strength of this component is specified. A
> comment should be added.

> - Sec. 2.1: More on this topic: do the fluid isobaric surfaces (i.e. the
> surfaces of constant fluid pressure) coincide with the magnetic
> isobaric surfaces (i.e. the surfaces of constant magnetic pressure)?
> How are these surfaces related to the iso-density surfaces? A comment
> should be added.
>

> - Sec. 3: the authors are a bit too sloppy when referring to the fluid
> pressure, which they call respectively: "fluid pressure", "gas
> pressure" "thermal pressure", and sometimes simply as "pressure". Since
> distinguishing the fluid, magnetic and total pressure is important, I
> ask the authors to always just talk of "fluid pressure".

We have followed the recomendation of the referee and in the revised version
of the paper only mentions to “fluid pressure”, “magnetic pressure”, and 
“total pressure” appear. Those appearances are highlighted in boldface
throughout.

> - Sec. 3: similar considerations apply to the density, which should
> be referred to as "rest-mass density"
>

Done.

> - Sec. 3: the authors discuss that the rest-mass density distribution can
> "be trivially obtained inverting the barotropic relation, p = K
> rho^kappa". I am not sure this is so trivial unless an assumption is
> made on the specific internal energy; the authors should discuss more
> clearly how they actually obtain the density from Eqs. (12) and (14).
>

> - Sec. 3: the authors discuss that Eq. (24) diverges for theta --> pi/2,
> i.e. at the equator. I am a bit puzzled by this statement and I wonder if
> this is an exact statement, ie whether it is true that

> partial_theta W = 0 for theta --> pi/2.

> If this is an exact result, I urge the authors to prove it. If it is
> not, I urge the authors to simply mention that the gradients become
> very small.
>

> - Fig. 1: as with the middle panel, I suggest that the right
> panel should also show the logarithm of the rest-mass density.
>

> - Fig. 2: all of the panels would benefit from a legend showing the
> corresponding values of gamma and beta. This would avoid having to read the caption repeatdly. The same is true also for Fig. 3.
>

> - Fig. 4: the caption should be corrected in "beta = 10^3 --> beta_{m_c}
> = 10^3".

> - Fig. 4: the authors do not discuss what is the largest magnetization
> that they can handle. I agree that beta_{m_c} = 10^{-3} is probably
> already quite high, but what happens for say beta_{m_c} = 10^{-6}? Can
> a solution still be found?

> - Sec. 4: I find the discussion in this section the less clear; a lot
> of the considerations appear phenomenological but I can't find the
> simple back-of-the-envelope demonstration that what the authors observe
> numerically is what should be expected on physical grounds. For
> example, the authors find that the rho_max increases with magnetization
> (left panel of Fig. 6) or that the position of the rest-mass density
> maximum moves in with increasing magnetization (middle panel of
> Fig. 6). What are the physical reasons for this behavior? I have some
> suggestions on this but I urge the authors to provide some simple and
> intuitive explanations first.
>

> - References: the list is already quite complete but maybe it can be
> complemented with the following ones, many of which are actually
> written by one of the authors.

> A comprehensive discussion of the properties of tori from merging binaries:

> http://adsabs.harvard.edu/abs/2016arXiv160703540B
>

> First detailed discussion of non-constant specific angular momentum tori:

> http://adsabs.harvard.edu/abs/2004MNRAS.354.1040M

> First discussion of the dynamics of magnetized tori built with the
> Komissarov solution

> http://adsabs.harvard.edu/abs/2007MNRAS.378.1101M

We have incorporated in the article the first and the third references 
suggested by the referee. However, we have not incorporated the second 
reference. While this is a very relevant work in the context of oscillations 
of non-constant-angular momentum disks, including this reference in our 
work seems to us a bit forced since it discusses aspects which are a bit 
away from the main discourse of our paper.

### 2017-06-16 paper_revised/paper_revised.tex
Revised version of the paper.
Work on the reply.

\documentclass{aa}

@@

Matter accretion on to black holes is the most efficient form of energy production known in nature. The conversion of the
gravitational energy of the infalling matter into heat and radiation may reach efficiencies of about 43\% in the case of maximally rotating (Kerr) black holes. For this reason, systems formed by a black hole surrounded by an accretion disc are deemed responsible for many of the most energetic astronomical phenomena observed in the cosmos. In particular, (geometrically) thick accretion discs (or tori) are believed to be present in quasars and other active galactic nuclei, some X-ray binaries and microquasars, as well as in the central engine of gamma-ray bursts.
The latter are alluded to in connection with mergers of neutron star binaries and black hole neutron star binaries, as well as with the rotational collapse that ensues at the end of the life of some massive stars. As numerical simulations show, such events  often result in a black hole surrounded by a torus (see e.g.~\cite{Rezzolla:2010,Sekiguchi:2011,Faber:2012,Shibata:2011}).

The investigation of this type of systems, either by analytical or numerical means, may rely on the ability to construct suitable 
representations based on physical assumptions. The construction of equilibrium models of stationary discs around black holes has indeed a long tradition (see~\cite{Abramowicz:2013} and references therein). In particular, the so-called ``Polish doughnuts"~\citep{Abramowicz:1978,Kozlowski:1978} provide a  very general method to build equilibrium configurations of perfect fluid matter orbiting around a Kerr black hole. This fully relativistic model assumes that the disc is non-magnetised and that the matter obeys a  constant specific angular momentum distribution. This method was later extended by~\cite{Komissarov:2006}  by adding a purely azimuthal magnetic field to build magnetised tori around rotating black holes. On the other hand, assuming different distributions of angular momentum in the discs, \citet{Qian:2009} presented a method to build sequences of black hole thick accretion discs in dynamical equilibria, restricted however to the purely hydrodynamical case.

@@

The equations of ideal general relativistic MHD are the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (w + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $w$ and $p$ being the fluid enthalpy and pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(w + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Following~\cite{Komissarov:2006} we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.
To integrate Eq.~\eqref{eq:diff_ver} we first assume a barotropic equation of state $w = w(p)$ of the form
\begin{equation}\label{eq:eos_fluid}
p = K w^{\kappa},
\end{equation}
with $K$ and $\kappa$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analog equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$~\citep{Komissarov:2006}
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = K_{\mathrm{m}} \tilde{w}_{\mathrm{m}}^{\lambda
},
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda
-1} w^{\lambda
},
\end{equation}
where $K_{\mathrm{m}}$ and $\lambda
$ are constants.
This particular choice of barotropic relationships, $w = w(p)$ and $\tilde{w} = \tilde{w}(\tilde{p}_{\mathrm{m}})$, fulfill the general relativistic version of the von Zeipel theorem for a toroidal magnetic field~\citep{vonZeipel:1924, Zanotti:2015}, i.e.~the surfaces of constant $\Omega$ and constant $l$ coincide.

@@

To construct our models of magnetised discs we follow the approach described in \citet{Qian:2009}. First, we find the radial location of the cusp and of the centre of the disc in the equatorial plane, $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$, defined as the solutions to the equation $l(r) - l_{\mathrm{K}} = 0$.
Next, we compute the partial derivatives of the potential, Eq.~\eqref{eq:potential}
\begin{equation}\label{eq:radial_der_pot}
\partial_r W = \partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\,,
\end{equation}
and
\begin{equation}\label{eq:polar_der_pot}
\partial_{\theta} W = \partial_{\theta} \ln|u_t| - \frac{\Omega \partial_{\theta}l}{1 - \Omega l}\,.
\end{equation}
Then, we integrate the radial partial derivative of the potential along the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ (assuming $W_{\mathrm{cusp}} = 0$) at the equatorial plane, thus obtaining the equatorial distribution of the potential between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$
\begin{equation}\label{eq:equatorial_pot}
W_{\mathrm{eq}}(r) = \int^{r_{\mathrm{c}}}_{r_{\mathrm{cusp}}}\left(\partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\right).
\end{equation}
Following~\citet{Qian:2009} (see also~\citet{Jaroszynski:1980}) we can divide equations \eqref{eq:radial_der_pot} and \eqref{eq:polar_der_pot} to obtain
\begin{equation}\label{eq:F}
F(r, \theta) = -\frac{\partial_r W}{\partial_{\theta} W} = \frac{\mathrm{d}\theta}{\mathrm{d}r}\,,
\end{equation}
where function $F(r, \theta)$ is known in closed form for the Kerr metric once an assumption for the angular momentum distribution has been made (cf.~Eq.~(\ref{eq:ansatz})). For this reason, Eq.~(\ref{eq:F}) also takes the form of an ordinary differential equation for the surfaces of constant potential, $\theta=\theta(r)$, which, upon integration, yields the location of those surfaces. Note that in~\citet{Qian:2009} these are surfaces of constant {\it pressure} instead, since their discs are purely hydrodynamical.

\begin{figure*}[t]
\centering
\includegraphics[scale=0.16]{figures/fig1a.pdf}
\hspace{-0.3cm}
\includegraphics[scale=0.16]{figures/fig1b.pdf}
\hspace{-0.2cm}
\includegraphics[scale=0.16]{figures/fig1c.pdf}
\caption{Comparison with Komissarov's solution for a constant angular momentum model. The left panel shows the density distribution in logarithm scale for all radii and all angles while the middle and right panels show, respectively, the angular profile of the logarithm of the density at the centre of the disc and the radial profile of the density at the equatorial plane.}
\label{komissarov}
\end{figure*}

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the gas pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the enthalpy at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the enthalpy distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$.

@@

\begin{table}
\caption{List of models for $\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{3}$. From left to right the columns report: the model name (with A, B and C standing for black hole spin $a = 0.5$, $0.9$ and $0.99$, respectively), the parameters of the angular momentum distribution $\beta$ and $\gamma$, the corresponding radii of the maximum thermal pressure and magnetic pressure, $r_{{\rm{max}}}$ and $r_{{\mathrm{m, max}}}$, the gravitational potential difference between the centre and the inner radius, $\Delta W\equiv W_{\rm c}-W_{\rm in}$, and the inner and outer radii of the discs, $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$.}             
\label{table:1}      
\centering          
\begin{tabular}{c c c c  c c c c}
\hline\hline       
 & $\beta$ & $\gamma$ & $r_{\rm{max}}$ &  $r_{\mathrm{m, max}}$ & $\Delta W$               & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ \\ 
 &              &                   &                          &                                        & $(\times 10^{-2})$     &                              &  \\
\hline           
A1 & $0.0$ & $0.0$ & $7.15$ &  $8.14$  & $-6.35$ & $2.91$ & -- \\ 
A2 & $0.5$ & $0.5$ & $7.15$ &  $7.65$  & $-2.27$ & $3.20$ & $11.8$\\ 
A3 & $0.9$ & $0.9$ & $7.15$ &  $7.45$  & $-0.30$ & $3.70$ &  $9.58$\\ 
A4 & $0.0$ & $0.9$ & $7.15$ &  $8.32$  & $-6.35$ & $2.91$ & $9.68$\\ 
A5 & $0.9$ & $0.0$ & $7.15$ &  $7.28$  & $-0.30$ & $3.70$ & --\\ 
 \hline 
B1 & $0.0$ & $0.0$ & $3.59$ &  $3.98$  & $-12.9$ & $1.73$ & -- \\ 
B2 & $0.5$ & $0.5$ & $3.59$ &  $3.83$  & $-4.32$ & $1.86$ & $5.35$\\ 
B3 & $0.9$ & $0.9$ & $3.59$ &  $3.73$  & $-0.54$ & $2.08$ & $4.52$\\ 
B4 & $0.0$ & $0.9$ & $3.59$ &  $3.85$  & $-12.9$ & $1.73$ & $4.54$\\ 
B5 & $0.9$ & $0.0$ & $3.59$ &  $3.81$  & $-0.54$ & $2.08$ & -- \\  
 \hline 
C1 & $0.0$ & $0.0$ & $1.98$ &  $2.28$  & $-24.6$ & $1.21$ & -- \\ 
C2 & $0.5$ & $0.5$ & $1.98$ &  $2.09$  & $-7.34$ & $1.26$ & $2.50$\\ 
C3 & $0.9$ & $0.9$ & $1.98$ &  $2.04$  & $-0.85$ & $1.35$ & $2.25$\\ 
C4 & $0.0$ & $0.9$ & $1.98$ &  $2.10$  & $-24.6$ & $1.21$ & $2.26$\\ 
C5 & $0.9$ & $0.0$ & $1.98$ &  $2.08$  & $-0.85$ & $1.35$ & --\\ 
\hline      
\end{tabular}
\end{table}

>>>>

\documentclass[referee]{aa}

@@

Matter accretion on to black holes is the most efficient form of energy production known in nature. The conversion of the
gravitational energy of the infalling matter into heat and radiation may reach efficiencies of about 43\% in the case of maximally rotating (Kerr) black holes. For this reason, systems formed by a black hole surrounded by an accretion disc are deemed responsible for many of the most energetic astronomical phenomena observed in the cosmos. In particular, (geometrically) thick accretion discs (or tori) are believed to be present in quasars and other active galactic nuclei, some X-ray binaries and microquasars, as well as in the central engine of gamma-ray bursts.
The latter are alluded to in connection with mergers of neutron star binaries and black hole neutron star binaries, as well as with the rotational collapse that ensues at the end of the life of some massive stars. As numerical simulations show, such events  often result in a black hole surrounded by a torus (see e.g.~\cite{Rezzolla:2010,Sekiguchi:2011,Faber:2012,Shibata:2011,Baiotti:2017}).

The investigation of this type of systems, either by analytical or numerical means, may rely on the ability to construct suitable 
representations based on physical assumptions. The construction of equilibrium models of stationary discs around black holes has indeed a long tradition (see~\cite{Abramowicz:2013} and references therein). In particular, the so-called ``Polish doughnuts"~\citep{Abramowicz:1978,Kozlowski:1978} provide a  very general method to build equilibrium configurations of perfect fluid matter orbiting around a Kerr black hole. This fully relativistic model assumes that the disc is non-magnetised and that the matter obeys a  constant specific angular momentum distribution. This method was later extended by~\cite{Komissarov:2006}  by adding a purely azimuthal magnetic field to build magnetised tori around rotating black holes. {\bf Dynamical evolutions of magnetized tori built with the Komissarov solution were first reported by~\cite{Montero:2007}.}
On the other hand, assuming different distributions of angular momentum in the discs, \citet{Qian:2009} presented a method to build sequences of black hole thick accretion discs in dynamical equilibria, restricted however to the purely hydrodynamical case.

@@

The equations of ideal general relativistic MHD are the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (w + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $w$ and $p$ being the fluid enthalpy and {\bf fluid} pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(w + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Following~\cite{Komissarov:2006} we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.
To integrate Eq.~\eqref{eq:diff_ver} we first assume a barotropic equation of state $w = w(p)$ of the form
\begin{equation}\label{eq:eos_fluid}
p = K w^{\kappa},
\end{equation}
with $K$ and $\kappa$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analog equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$~\citep{Komissarov:2006}
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = K_{\mathrm{m}} \tilde{w}_{\mathrm{m}}^{\lambda
},
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda
-1} w^{\lambda
},
\end{equation}
where $K_{\mathrm{m}}$ and $\lambda
$ are constants.
This particular choice of barotropic relationships, $w = w(p)$ and $\tilde{w} = \tilde{w}(\tilde{p}_{\mathrm{m}})$, fulfill the general relativistic version of the von Zeipel theorem for a toroidal magnetic field~\citep{vonZeipel:1924, Zanotti:2015}, i.e.~the surfaces of constant $\Omega$ and constant $l$ coincide.

@@

To construct our models of magnetised discs we follow the approach described in \citet{Qian:2009}. First, we find the radial location of the cusp and of the centre of the disc in the equatorial plane, $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$, defined as the solutions to the equation $l(r) - l_{\mathrm{K}} = 0$.
Next, we compute the partial derivatives of the potential, Eq.~\eqref{eq:potential}
\begin{equation}\label{eq:radial_der_pot}
\partial_r W = \partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\,,
\end{equation}
and
\begin{equation}\label{eq:polar_der_pot}
\partial_{\theta} W = \partial_{\theta} \ln|u_t| - \frac{\Omega \partial_{\theta}l}{1 - \Omega l}\,.
\end{equation}
Then, we integrate the radial partial derivative of the potential along the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ (assuming $W_{\mathrm{cusp}} = 0$) at the equatorial plane, thus obtaining the equatorial distribution of the potential between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$
\begin{equation}\label{eq:equatorial_pot}
W_{\mathrm{eq}}(r) = \int^{r_{\mathrm{c}}}_{r_{\mathrm{cusp}}}\left(\partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\right).
\end{equation}
Following~\citet{Qian:2009} (see also~\citet{Jaroszynski:1980}) we can divide equations \eqref{eq:radial_der_pot} and \eqref{eq:polar_der_pot} to obtain
\begin{equation}\label{eq:F}
F(r, \theta) = -\frac{\partial_r W}{\partial_{\theta} W} = \frac{\mathrm{d}\theta}{\mathrm{d}r}\,,
\end{equation}
where function $F(r, \theta)$ is known in closed form for the Kerr metric once an assumption for the angular momentum distribution has been made (cf.~Eq.~(\ref{eq:ansatz})). For this reason, Eq.~(\ref{eq:F}) also takes the form of an ordinary differential equation for the surfaces of constant potential, $\theta=\theta(r)$, which, upon integration, yields the location of those surfaces. Note that in~\citet{Qian:2009} these are surfaces of constant {\it {\bf fluid} pressure} instead, since their discs are purely hydrodynamical.

\begin{figure*}[t]
\centering
\includegraphics[scale=0.14]{figures/fig1a.pdf}
\hspace{-0.3cm}
\includegraphics[scale=0.14]{figures/fig1b.pdf}
\hspace{-0.2cm}
\includegraphics[scale=0.14]{figures/fig1c.pdf}
\caption{Comparison with Komissarov's solution for a constant angular momentum model. The left panel shows the {\bf rest-mass} density distribution in logarithm scale for all radii and all angles while the middle and right panels show, respectively, the angular profile of the logarithm of the {\bf rest-mass} density at the centre of the disc and the radial profile of the {\bf rest-mass} density at the equatorial plane.}
\label{komissarov}
\end{figure*}

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the {\bf fluid} pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the enthalpy at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the enthalpy distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the {\bf rest-mass} density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$.

@@

\begin{table}
\caption{List of models for $\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{3}$. From left to right the columns report: the model name (with A, B and C standing for black hole spin $a = 0.5$, $0.9$ and $0.99$, respectively), the parameters of the angular momentum distribution $\beta$ and $\gamma$, the corresponding radii of the maximum {\bf fluid} pressure and magnetic pressure, $r_{{\rm{max}}}$ and $r_{{\mathrm{m, max}}}$, the gravitational potential difference between the centre and the inner radius, $\Delta W\equiv W_{\rm c}-W_{\rm in}$, and the inner and outer radii of the discs, $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$.}             
\label{table:1}      
\centering          
\begin{tabular}{c c c c  c c c c}
\hline\hline       
 & $\beta$ & $\gamma$ & $r_{\rm{max}}$ &  $r_{\mathrm{m, max}}$ & $\Delta W$               & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ \\ 
 &              &                   &                          &                                        & $(\times 10^{-2})$     &                              &  \\
\hline           
A1 & $0.0$ & $0.0$ & $7.15$ &  $8.14$  & $-6.35$ & $2.91$ & -- \\ 
A2 & $0.5$ & $0.5$ & $7.15$ &  $7.65$  & $-2.27$ & $3.20$ & $11.8$\\ 
A3 & $0.9$ & $0.9$ & $7.15$ &  $7.45$  & $-0.30$ & $3.70$ &  $9.58$\\ 
A4 & $0.0$ & $0.9$ & $7.15$ &  $8.32$  & $-6.35$ & $2.91$ & $9.68$\\ 
A5 & $0.9$ & $0.0$ & $7.15$ &  $7.28$  & $-0.30$ & $3.70$ & --\\ 
 \hline 
B1 & $0.0$ & $0.0$ & $3.59$ &  $3.98$  & $-12.9$ & $1.73$ & -- \\ 
B2 & $0.5$ & $0.5$ & $3.59$ &  $3.83$  & $-4.32$ & $1.86$ & $5.35$\\ 
B3 & $0.9$ & $0.9$ & $3.59$ &  $3.73$  & $-0.54$ & $2.08$ & $4.52$\\ 
B4 & $0.0$ & $0.9$ & $3.59$ &  $3.85$  & $-12.9$ & $1.73$ & $4.54$\\ 
B5 & $0.9$ & $0.0$ & $3.59$ &  $3.81$  & $-0.54$ & $2.08$ & -- \\  
 \hline 
C1 & $0.0$ & $0.0$ & $1.98$ &  $2.28$  & $-24.6$ & $1.21$ & -- \\ 
C2 & $0.5$ & $0.5$ & $1.98$ &  $2.09$  & $-7.34$ & $1.26$ & $2.50$\\ 
C3 & $0.9$ & $0.9$ & $1.98$ &  $2.04$  & $-0.85$ & $1.35$ & $2.25$\\ 
C4 & $0.0$ & $0.9$ & $1.98$ &  $2.10$  & $-24.6$ & $1.21$ & $2.26$\\ 
C5 & $0.9$ & $0.0$ & $1.98$ &  $2.08$  & $-0.85$ & $1.35$ & --\\ 
\hline      
\end{tabular}
\end{table}
